---
title: "ltvProject"
output: html_document
---


## R Markdown

Our client is an online greeting card company. The company offers monthly subscriptions at a rate of $1 per month for access to their eCard website. The client is interested in understanding the life-time value (ltv) of their customers.
The life-time value of a customer is defined as the total revenue earned by the company over the course of their relationship with the customer.
The enclosed (synthetic) data represent usage statistics for 10,000 customers. Usage is summarized at a daily level and covers a period of 4 years from 2011-01-01 to 2014-12-31.
The following is a description of each field captured in the enclosed data set containing a total of 10,000 customers.

| Data Field | Description                                                                         | 
|------------|-------------------------------------------------------------------------------------|
| `id`       | A unique user identifier                                                            |
| `status`   | Subscription status ‘0’- new, ‘1’- open, ‘2’- cancelation event                     |
| `gender`   | User gender ‘M’- male, ‘F’- female                                                  |
| `date`     | Date of in which user ‘id’ logged into the site                                     |
| `pages`    | Number of pages visted by user ‘id’ on date ‘date’                                  |
| `onsite`   | Number of minutes spent on site by user ‘id’ on date ‘date’                         |
| `entered`  | Flag indicating whether or not user entered the send order path on date ‘date’      |
| `completed`| Flag indicating whether the user completed the order (sent an eCard)                |
| `holiday`  | Flag indicating whether at least one completed order included a holiday themed card |

We must preprocess the data to determine the following: 

| Data Field | Description                                                                         | 
|------------|-------------------------------------------------------------------------------------|
| `lifespan` | The lifespan of each customer in days, for cancelled customers= cancelDate-openDate,|
|            | for open customers=maxDate-openDate+1/(cancelled customers/total customers)         |


```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
library(ISLR)
library(partykit)
library(caret)
library(rpart)
library(randomForest)
library(pROC)
library(readxl)
library(binaryLogic)
```

```{r, cache = TRUE}
# Importing teh data directly from Excel
customer.data <- read_excel("Downloads/ltv Dataset.xlsx", sheet = "Sheet1")
```

```{r, cache = TRUE}
# Transform the data to teh desired format
customer.data <- transform(
  customer.data,
  id=as.integer(id),
  status=as.integer(status),
  gender=as.factor(gender),
  date=as.Date(date),
  pages=as.integer(pages),
  onsite=as.integer(pages),
  entered=as.integer(entered),
  completed=as.integer(completed),
  holiday=as.integer(holiday)
)
```

```{r}
#processing the date
customer.data$date <- as.Date(customer.data$date,'%m/%d/%Y')
customer.data$month <- months(customer.data$date)
customer.data$year <- format(customer.data$date,format = '%Y')
```

```{r}
#calculate the average value
ltv.pages <- aggregate( pages ~ id+ month + year, customer.data, mean)
ltv.onsite <- aggregate( onsite ~ id + month + year, customer.data, mean)
#merger the table
ltv.m1 <- merge(x = ltv.pages, y = ltv.onsite, by = c('id','month','year'), all.x = TRUE)
```

```{r}
#sort the dataframe and export it
ltv.m1 <- ltv.m1[order(ltv.m1$id),]
ltv.m1
write.csv(ltv.m1,'ltv_modelOne.csv',row.names = FALSE)
```

```{r, cache=TRUE}
#convert the data from numeric to date type (duplicate) 
#customer.data$date <- as.Date(customer.data$date, origin = "1899-12-30")
#Calculate the customer lifespan 
#first group the data by ID to find the max and min date for a given customer
#along with the latest status 
customer.lifespan <- customer.data[, c("id", "date", "status")] %>% group_by(id)
customer.lifespan <-customer.lifespan %>% mutate(maxDate = max(date)) 
customer.lifespan <-customer.lifespan %>% mutate(status = max(status)) 

# there is a little bug in this code (need to be fixed)
customer.lifespan <- customer.lifespan %>% filter(date == min(date)) %>% rename(minDate = date)

#Subtract the maxDate and minDate to determine the number of days of subscription
customer.lifespan$subDays <- difftime(customer.lifespan$maxDate, customer.lifespan$minDate, units = "days")
#Determine the observed lifespan factor to be added
lifespanFraction <- 1/(with(1, sum(customer.data$status == 2))/10000)
#calculate the lifespan for the customers
customer.lifespan$lifespan <- ifelse(customer.lifespan$status == 2, customer.lifespan$subDays, customer.lifespan$subDays + lifespanFraction)
#add this data to the main dataset 
customer.data$lifespan <- customer.lifespan$lifespan[match(customer.data$id,customer.lifespan$id)]
```

```{r, cache = TRUE}
# Add 2 columns in the dataframe representing completed/holiday and onsite/entered
ltv_afterProcess <- transform(
  customer.data,
  CompletedVSHoliday=as.integer(entered)/as.integer(holiday),
  OnsiteVSEntered=as.integer(onsite)/as.integer(entered)
)
```

```{r, cache = TRUE}
# calculate the ratio between sum of all entered and sum of all completed
SumEnteredVSCompleted <- sum(ltv_afterProcess$entered)/sum(ltv_afterProcess$completed)
SumEnteredVSCompleted
```

```{r, cache = TRUE}
# create a new dataframe representing teh aggregated summation of each variable per customer
aggregatedCustomerSums <- aggregate(cbind(PagesSum=ltv_afterProcess$pages, OnsiteSum=ltv_afterProcess$onsite, EnteredSum=ltv_afterProcess$entered, CompletedSum=ltv_afterProcess$completed, HolidaySum=ltv_afterProcess$holiday), by=list(Customerid=ltv_afterProcess$id), FUN=sum)
#aggregatedCustomerSums <- transform(
  #aggregatedCustomerSums,
  # a little bug in line 79 need to be fixed
  #Lifespan=customer.lifespan[, "lifespan"] %>% group_by(customer.lifespan$id)
  #Gender=customer.lifespan$gender
#)
```

```{r}
#pre-processed dataset outputs
view(ltv.m1)
view(customer.lifespan)
view(aggregatedCustomerSums)
```


## 1.	Develop an attrition model, to predict whether a customer will cancel their subscription in the near future. Characterize your model performance.

```{r}
```

## 2.	Develop a model for estimating the ltv of a customer. Characterize your model performance.

```{r}
```

## 3.	Develop a customer segmentation scheme. Include in this scheme the identification of sleeping customers, those that are no longer active but have not canceled their account.

```{r}
```
